---
title: "Can we predict house prices using known features of each house and a supervised learning approach?"
author: "Florence Galliers"
date: "18/10/2020"
output: pdf_document
---
```{r set wd / libraries / reading in data, include = FALSE}
setwd("~/Google Drive/Harper/1-C7081/Asssesment/github-C7081/C7081-assessment")

library()

data <- read.csv(file = "original-data.csv")
```

## Background:
House prices are an important part of the economy...
Motivate the story...
What do we already know?
What lessons will the story teach us?

When purchasing a house people are more likely to look only at the number of bedrooms and number of bathrooms, rather than the square foot of the house. This may be because they do not understand how square foot is calculated, or find it difficult to visualise how this translates into the house size. 

### Objectives:
- Understand which independent variables in the data set can be used to predict house price (the dependent variable)
- Minimize the differences between predicted and actual house prices by using model selection to choose the most accurate model.

## Data:
The data for this analysis contains house prices and information regarding the features of each house. It has 4600 entries. The original data set downloaded from Kaggle had 17 independent variables, however I felt that 7 of these were not relevant to this analysis and so they have been removed. The dependent variable in this analysis will be the house price (in US dollars). 

Although it is from 2014 and from houses in the USA, I feel it is an interesting data set that contains a large amount of information and number of variables. I was not able to find a similar data set from the UK, and so I am going to go ahead with this data set as I feel it will produce some interesting results.

### Data Preparation:
- There was originally a variable called sqft_basement which gave the size of the basement if present, 
however a lot of houses did not have basement so I felt it would be more useful to turn this variable from numeric 
into binary. It now shows 0 = no basement, 1 = basement. 
- I also changed the variable year_renovated into a binary variable, as again, not all the houses had been renovated. 0 = not renovated, 1 = renovated.
- Check if there are any zero values for the price variable as these are not acceptable in a housing price situation, a house cannot cost nothing, so we must remove these and assume they are errors in the data set.
- Remove outliers

```{r, echo = FALSE}
# removing unwanted variables from original kaggle dataset
variables_to_remove <- c(1, 8, 9, 11, 15, 17, 18)
data <- data[ ,-variables_to_remove]

# changing variable sqft_basement into binary variable basement
data$sqft_basement <- ifelse(data$sqft_basement == 0, 0, 1)
# change column name from sqft_basement to if_basement
names(data)[names(data) =="sqft_basement"] <- "if_basement"

# changing yr_renovated into binary variable renovates
data$yr_renovated <- ifelse(data$yr_renovated == 0, 0, 1)
# change column name from yr_renovated to if_renovated
names(data)[names(data) =="yr_renovated"] <- "if_renovated"

# remove zero values for house price
```


## Methods:
As house price is a continuous variable I have taken a supervised learning approach and will be using regression to look at the relationship between house price and features of each house.

Review approaches tried or considered

Summary of final approach and justification of why this approach was chosen:


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```


## Results:

Summary of major results, graphs, diagnostic outputs

Strictly relevant to the objectives

Must include a link to the Github repository containing a fully reproducible and documented analysis
Reported in scientific style.



## Conclusions:

~1 paragraph


### Literature Cited:
3-5 peer reviewed references
-
-
-
-
-