---
title: "Can we predict house prices using known features of each house and a supervised learning approach?"
author: "Florence Galliers"
date: "18/10/2020"
output: pdf_document
---
```{r Reading in Data, include = FALSE}
library(openxlsx) 
data <- read.xlsx("housing_data_assessment.xlsx")
```

## Background:
House prices are an important part of the economy...

### Objectives:
- Understand which independent variables in the data set can be used to predict house price (the dependent variable)
- Minimize the differences between predicted and actual house prices by using 
model selection to choose the most accurate model.

## Data:
The data used for this analysis contains house prices and information regarding 
the house features. It has 4600 entries. The original data set downloaded from 
Kaggle had 17 independent variables, however I felt that 7 of these were 
not relevant to this analysis and so they have been removed. The dependent 
variable in this analysis will be the house price (in US dollars). 


### Data Preparation:
- changed 2 variables to binary, talk about this here
- dummy variables for condition variable??
- prices are large numbers, change to show in thousands rather than pounds?

```{r, echo = FALSE}
head(data)
```



## Methods:
As house price is a continuous variable I have taken a supervised learning approach
and will be using regression to look at the relationship between house price
and features of each house.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```


## Results:


## Conclusions:



### Literature Cited: